---
title: "EQ_Tsunami_Analysis"
author: "Beyza Özen"
date: "12/20/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
Data source: https://www.kaggle.com/kanncaa1/earthquake-animation-with-plotly/notebook
```{r}
library(readr)
earthquake <- read_csv("data/earthquake.csv")
```

```{r}
library(dplyr)
glimpse(earthquake)
```


```{r}
library(lubridate)
earthquake$Date <- dmy(earthquake$Date)
```
# Earthquake magnitude-depth relation
In this part, I inspired from the article given below. 
https://www.researchgate.net/figure/A-scatter-plot-for-M-W-versus-depth-km-for-the-earthquakes-in-the-F-net-catalog_fig11_283867433

```{r}
library(ggplot2)


p<-earthquake %>% 
  filter(!is.na(Magnitude) & !is.na(Depth) & Depth != 0)%>% 
  ggplot()+
  geom_point(aes(x = Depth, y = Magnitude)) 

ggExtra::ggMarginal(p, type = "densigram")

```

This plot is easy with ggMarginal function provided by package ggExtra. I want to show is distributions for depth and magnitude of earthquakes. I know that magnitude distribution has a exponential curve. This is why I want to add logarithmic scale to the histograms. Unfortunately I could not find a mthod on ggMarginal to make it happen. So I use the code below to create logarithmic y scales for frequencies. 


```{r}
hist_top <- ggplot()+
  geom_histogram(aes(earthquake$Depth))+
  scale_y_continuous(name = "" , trans='log10')+
  theme(axis.title.x=element_blank(), axis.text.x=element_blank())


empty <- ggplot()+geom_point(aes(1,1), colour="white")+
         theme(axis.ticks=element_blank(), 
               panel.background=element_blank(), 
               axis.text.x=element_blank(), axis.text.y=element_blank(),           
               axis.title.x=element_blank(), axis.title.y=element_blank())


hist_right <- ggplot()+
  geom_histogram(aes(earthquake$Magnitude))+
  scale_y_continuous(name = "" , trans='log10')+
  coord_flip()+
  theme(axis.title.y=element_blank(), axis.text.y=element_blank())

library(gridExtra)
grid.arrange(hist_top, empty, p, hist_right, ncol=2, nrow=2, widths=c(4, 1), heights=c(1, 4))
```


# Locational earthquake visualisation with a small inner map focused on Japan

World map coordinates are avaiblable on tidyverse
```{r}
world <- map_data("world")
```
filtering the coordinate data of Japan and it subregions into new variable. 
```{r}
japan <- world %>% filter(region== 'Japan') 
```

Filtering earthquake data so that I can only focus on earthquakes larger than magnitude 7. Also I will only consider MW as magnitude type since it is the most stable one for higher magnitudes. 
```{r}
earthquake7 <- earthquake %>% filter(Magnitude>=7 & `Magnitude Type` =='MW')
```
To visualization, I referenced from the below 2 websites. I combined the 2 method for my data.
  - https://datavizpyr.com/how-to-make-world-map-with-ggplot2-in-r/
  - https://datavizpyr.com/how-to-zoom-in-on-a-plot-in-r/
```{r}
library(ggforce)
ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
        color = "black", fill = "lightgray", size = 0.2
  ) +
  geom_point(
    data = earthquake7,
    aes(Longitude, Latitude, alpha= Magnitude, size = Magnitude),
  ) +
  ggtitle("Japan") +
  facet_zoom(ylim = c(min(japan$lat), max(japan$lat)), xlim = c(min(japan$long), max(japan$long)),zoom.size = 0.4)
```


# Depth vs magnitude with locational seperations 
In this last part, I would like to focus on the animation given in my data source. The person who upload the data that I use on to Kaggle also provide a beautiful animation with time. S/he used to Python. My purpose is to recreate it using shiny on r. 
I use the webside below as reference to create the visualisation.
  https://bookdown.org/ihaddaden_fodeil/shiny_book-master/reactivity.html
```{r}
eq <- earthquake %>% filter(!is.na(Date))
eq$year <- as.numeric(format(eq$Date, format="%Y"))
```

```{r}
library(shiny)

ui <- fluidPage(

  sliderInput(inputId = "Date", 
            label = "Date",
            min = min(eq$year), max = max(eq$year),value = 2001,step=1), 
  
  plotOutput(outputId = "plot")
  
)

server <- function(input, output) {
  
  
output$plot <- renderPlot({
  
  ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region),
        color = "white", fill = "lightgray", size = 0.2
  ) +
  geom_point(
    data = eq %>% filter(year==input$Date),
    aes(Longitude, Latitude, size = Magnitude,color =Depth),
    alpha = 0.6
  ) 
  
})
  
}

shinyApp(ui = ui, server = server)
```



